Planning to Save Energy: How Information Format Affects Accuracy

### 2025-01-09 manuscript/appendix.qmd
todays  updates - intro, discussion, few methods tweaks

# A tibble: 4,560 × 17
# Groups:   id, block [380]
   id     block refClass appliance pct_goal rounded state    calc       edu                         els   pct family state_avg plan  value change state_dif
   <chr>  <dbl> <fct>    <chr>     <chr>    <chr>   <chr>    <chr>      <fct>                     <dbl> <dbl>  <dbl>     <dbl> <fct> <dbl>  <dbl>     <dbl>
 1 s2_100     1 kWh      AC        10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000       498 plan1   500   -500         2
 2 s2_100     1 kWh      AC        10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000       498 plan2   300   -700      -198
 3 s2_100     1 kWh      HEAT      10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  18000     16411 plan1 15000  -3000     -1411
 4 s2_100     1 kWh      HEAT      10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  18000     16411 plan2 12000  -6000     -4411
 5 s2_100     1 kWh      WATER     10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  12000      5832 plan1  6000  -6000       168
 6 s2_100     1 kWh      WATER     10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1  12000      5832 plan2  2000 -10000     -3832
 7 s2_100     1 kWh      REFRIG    10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000      1142 plan1  2000   1000       858
 8 s2_100     1 kWh      REFRIG    10%      Rounded Colorado Calculator Highschool diploma or GED 0.875   0.1   1000      1142 plan2   500   -500      -642

```{r}
#| eval: false
# widen by appliance, with separate row for value, family, and stage_avg

s2 <- s2_long #|> select(-family,-state_avg,-edu,-change,-state_dif) |> pivot_wider(names_from = appliance, values_from = c(value)) 
s2s <- s2_long |> select(-family,-state_avg,-edu,-change,-value) |> pivot_wider(names_from = appliance, values_from = c(state_dif)) |> select(id,block,refClass,pct,plan,AC:OTHER)
# s2s
#    id     block refClass     pct plan     AC   HEAT WATER REFRIG OTHER
#    <chr>  <dbl> <fct>      <dbl> <fct> <dbl>  <dbl> <dbl>  <dbl> <dbl>
#  1 s2_100     1 kWh         0.1  plan1     2  -1411   168    858   348
#  2 s2_100     1 kWh         0.1  plan2  -198  -4411 -3832   -642 -6152
#  3 s2_100     2 kWh         0.15 plan1  1324     19  -139    320  -424
#  4 s2_100     2 kWh         0.15 plan2   324   -981 -1139   -318 -1424
#  5 s2_101     1 kWh         0.1  plan1   102   -411  2168   -142   398
#  6 s2_101     1 kWh         0.1  plan2   502  -1411   168   -142  1348
#  7 s2_101     2 kWh         0.15 plan1   751    901  -396    682 -1883
#  8 s2_101     2 kWh         0.15 plan2  2251    401   104    682 -1383
#  9 s2_102     1 Percentage  0.1  plan1   958 -12954 -5354    200 -5417
# 10 s2_102     1 Percentage  0.1  plan2   890 -15435 -4956   8544  3225

# clustering planning strategies - kmeans

s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) 
s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor() 
s2s <- cbind(s2s,c=s2s_clust)

s2s |>    group_by(c) |>
  summarise(ac=mean(AC),heat=mean(HEAT),water=mean(WATER),refrig=mean(REFRIG),other=mean(OTHER)) |>
  pivot_longer(cols = c(ac:other), names_to = "Appliance", values_to = "Mean") |>
  ggplot(aes(x=Appliance,y=Mean,col=c)) + 
  geom_point() + 
  facet_wrap(~c) +
  theme_minimal()

s2s <- s2_long |> select(-family,-state_avg,-edu,-change,-state_dif) |> pivot_wider(names_from = appliance, values_from = c(value))  |> select(id,block,refClass,pct,plan,AC:OTHER)

s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3)

s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor()

s2s <- cbind(s2s,c=s2s_clust) |> left_join(s2, join_by(id,refClass,block,plan,pct))
s2s |>  group_by(c) |>
  summarise(ac=mean(AC),heat=mean(HEAT),water=mean(WATER),refrig=mean(REFRIG),other=mean(OTHER)) |>
  pivot_longer(cols = c(ac:other), names_to = "Appliance", values_to = "Mean") |>
  ggplot(aes(x=Appliance,y=Mean,col=c)) + 
  geom_point() + 
  geom_point(aes(y=state_avg),color="black") +
  facet_wrap(~c) +
  theme_minimal()

s2 |> group_by(appliance,family,state,state_avg) |> select(appliance,family_prior_yr=family,state,state_avg)
```

@@

n_exclusion <-readRDS(here::here("data/s1_grp_outlier_ids.rds")) |> length() # 6

@@

gender_counts <- s1 |> 
  group_by(id,gender) |>
  slice(1) |>
  ungroup() |> group_by(gender) |> 
  summarise(count = n()) |>
  ungroup()

@@

@@

n_exclusion <-readRDS(here::here("data/s2_grp_outlier_ids.rds")) |> length() # 6

@@

gender_counts <- s2_long |> 
  group_by(id,gender) |> 
  slice(1) |>
  ungroup() |> group_by(gender) |>
  summarise(count = n()) |>
  ungroup()
#   gender count
#   <fct>  <int>
# 1 Male     102
# 2 Female    88

>>>>

# Subject Counts

s1 %>%
  group_by(id) %>%
  summarise(total_subjects = n_distinct(id)) %>%
  ungroup() %>%
  summarise(total_subjects = n_distinct(id)) %>%
  bind_cols(
    s1 %>%
      group_by(refClass) %>%
      summarise(subjects_per_refClass = n_distinct(id)) %>%
      ungroup()
  ) %>%
  print()

s2_long %>%
  group_by(id) %>%
  summarise(total_subjects = n_distinct(id)) %>%
  ungroup() %>%
  summarise(total_subjects = n_distinct(id)) %>%
  bind_cols(
    s2_long %>%
      group_by(refClass) %>%
      summarise(subjects_per_refClass = n_distinct(id)) %>%
      ungroup()
  ) %>%
  print()

@@

n_exclusion_s1 <-readRDS(here::here("data/s1_grp_outlier_ids.rds")) |> length() # 6

@@

gender_counts <- s1 |> 
  group_by(id,gender) |>
  slice(1) |>
  group_by(gender) |> 
  summarise(count = n())

@@

s1 %>% ungroup() |> 
  summarise(
    total_subjects = n_distinct(id),
    .groups = 'drop'
  ) %>%
  left_join(
    s1 %>%
      group_by(refClass) %>%
      summarise(
        subjects_in_class = n_distinct(id),
        .groups = 'drop'
      )
  ) %>%
  print()

 s1 %>% summarise(
      total_subjects = n(),
      reference_class_counts = list(
        distinct(., id, !!sym(refClass)) %>% count(!!sym(refClass))
      )
    )

@@

n_exclusion_s2 <-readRDS(here::here("data/s2_grp_outlier_ids.rds")) |> length() # 6

@@

gender_counts <- s2_long |> 
  group_by(id,gender) |> 
  slice(1) |>
  group_by(gender) |>
  summarise(count = n()) 
#   gender count
#   <fct>  <int>
# 1 Male     102
# 2 Female    88
